The perception of places and the construction of territorial diagnostics are increasingly mediated by digital technologies and by data spontaneously produced by citizens. This dissertation investigates how advanced sentiment analysis, supported by Large Language Models (LLMs), can improve the interpretation of these subjective experiences and transform them into useful information for monitoring Sustainable Development Goals (SDGs) at the municipal level. The research adopts the Design Science methodology to guide the conception, implementation, and evaluation of a computational artifact capable of collecting georeferenced public reviews, applying geographic and semantic relevance filters, performing aspect-based sentiment analysis, and generating structured diagnostics in natural language. The approach uses data from the Google Places API and models such as Gemini and Perplexity to analyze perceptions associated with services, institutions, urban infrastructure, and other places related to the SDGs. The main objective is to answer the following question: how can aspect-based sentiment analysis, applied to georeferenced perceptual data, help identify bottlenecks in municipal SDG performance and support public policy? The application in the context of Mossoró-RN demonstrates the potential of the methodology to reveal recurring patterns in public comments, identify critical dimensions across different SDGs, support the construction of interactive dashboards, and provide actionable insights for public managers without requiring advanced technical expertise. Thus, this research aims to validate a scalable method for capturing, interpreting, and transforming citizens' perceptions of places into actionable diagnostics for urban planning, public management, and sustainable development.

% Separe as Keywords por ponto
\keywords{Sustainable Development Goals. Place Perception. Natural Language Processing. Sentiment Analysis. Large Language Models. Google Places API. Public Policy.}
